---
title: "Estimating nlmixr2 models with 'nlmixr2targets'"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Estimating nlmixr2 models with 'nlmixr2targets'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nlmixr2targets)
```

# Introduction to `nlmixr2targets`

The `nlmixr2targets` improves reproducibility by ensuring that your model is
up-to-date with your data, and it speeds your workflow using the `targets`
package to only run models when the model or data have changed.

There are two main functions that are used within the package:

* `tar_nlmixr()` which runs a single model, and
* `tar_nlmixr_multimodel()` which runs multiple models for a single dataset.

Using `nlmixr2targets` requires the use of the `targets` package.  To learn
about the `targets` package, see
(https://docs.ropensci.org/targets/)[the targets website].

# Model modifications to use `nlmixr2targets`

When running a model with `nlmixr2targets`, it must be interpretable by the
`targets` package.  The only notable issue comes from setting initial conditions
for a compartment, such as using `pd(0) <- initialConc`.  The reason this
doesn't work is the `targets` package does not allow the assignment into zero.
So, for any initial conditions, you will need to set them with `initial` instead
of `0`.  The example above would be `pd(initial) <- initialConc`, and
`nlmixr2targets` will convert it to the model `nlmixr2` expects for you.

# Running one model with one dataset (`tar_nlmixr()`)

The `tar_nlmixr()` function allows you to estimate one model with one dataset.
It will generate three targets:  a simplified version of the model, a minimal
version of the dataset, and the estimation step.

The simplified version of the model removes parts that are less reproducible but
changes none of the model intent.  (Advanced information:  The parts that are
removed are that the source references and the model name.  Also, the model is
modified at this step for setting initial values as described in the previous
section of this vignette.)

```{r using-tar_nlmixr, eval = FALSE}
library(targets)
library(tarchetypes)
library(nlmixr2targets)

pheno <- function() {
  ini({
    lcl <- log(0.008); label("Typical value of clearance")
    lvc <-  log(0.6); label("Typical value of volume of distribution")
    etalcl + etalvc ~ c(1,
                        0.01, 1)
    cpaddSd <- 0.1; label("residual variability")
  })
  model({
    cl <- exp(lcl + etalcl)
    vc <- exp(lvc + etalvc)
    kel <- cl/vc
    d/dt(central) <- -kel*central
    cp <- central/vc
    cp ~ add(cpaddSd)
  })
}

plan_model <-
  tar_plan(
    myData = nlmixr2data::pheno_sd,
    tar_nlmixr(
      model_pheno,
      object = pheno,
      data = myData,
      est = "saem"
    )
  )

list(
  plan_model
)
```

# Running multiple models with one dataset (`tar_nlmixr_multimodel()`)

A common use case is to generate multiple models using a single dataset and
estimation method.  `tar_nlmixr_multimodel()` allows the generation of a named
list of models to allow subsequent analysis of all models.

Internally, `tar_nlmixr_multimodel()` passes all the models to `tar_nlmixr()` so
that the data set simplification and equivalent steps run once per model, and
not model is run more often than required for dataset or model changes.

```{r using-tar_nlmixr_multimodel, eval = FALSE}
library(targets)
library(tarchetypes)
library(nlmixr2targets)

pheno <- function() {
  ini({
    lcl <- log(0.008); label("Typical value of clearance")
    lvc <-  log(0.6); label("Typical value of volume of distribution")
    etalcl + etalvc ~ c(1,
                        0.01, 1)
    cpaddSd <- 0.1; label("residual variability")
  })
  model({
    cl <- exp(lcl + etalcl)
    vc <- exp(lvc + etalvc)
    kel <- cl/vc
    d/dt(central) <- -kel*central
    cp <- central/vc
    cp ~ add(cpaddSd)
  })
}

pheno2 <- function() {
  ini({
    lcl <- log(0.008); label("Typical value of clearance")
    lvc <-  log(0.6); label("Typical value of volume of distribution")
    etalcl + etalvc ~ c(2,
                        0.01, 2)
    cpaddSd <- 3.0; label("residual variability")
  })
  model({
    cl <- exp(lcl + etalcl)
    vc <- exp(lvc + etalvc)
    kel <- cl/vc
    d/dt(central) <- -kel*central
    cp <- central/vc
    cp ~ add(cpaddSd)
  })
}

plan_model <-
  tar_nlmixr_multimodel(
    all_models,
    data = nlmixr2data::pheno_sd,
    est = "saem",
    "Base one-compartment model; IIV in clearance and volume; additive residual error" = pheno,
    "Base one-compartment model; IIV in clearance and volume; additive residual error (estimate starting at 3)" = pheno2
  )

plan_report <-
  tar_plan(
    # Determine the AIC for all tested models
    aic_list = sapply(X = all_models, FUN = AIC)
  )

list(
  plan_model,
  plan_report
)
```
