---
title: "Estimating nlmixr2 models with 'nlmixr2targets'"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Estimating nlmixr2 models with 'nlmixr2targets'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nlmixr2targets)
```

# Introduction to `nlmixr2targets`

The `nlmixr2targets` improves reproducibility by ensuring that your model is
up-to-date with your data, and it speeds your workflow using the `targets`
package to only run models when the model or data have changed.

There are two main functions that are used within the package:

* `tar_nlmixr()` which runs a single model, and
* `tar_nlmixr_multimodel()` which runs multiple models for a single dataset.

Using `nlmixr2targets` requires the use of the `targets` package.  To learn
about the `targets` package, see
(https://docs.ropensci.org/targets/)[the targets website].

# `tar_nlmixr()` usage
