---
title: "Events for loggers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Events for loggers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
has_bus <- exists("rxEventListen", asNamespace("rxode2"))
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = has_bus
)
```

```{r, include = FALSE, eval = TRUE}
note <- if (has_bus) {
  ""
} else {
  paste0(
    "> **Note:** the installed rxode2 has no event bus, so the code in this ",
    "vignette is shown but not run."
  )
}
```

`r note`

```{r setup}
library(nlmixr2targets)
```

# The rxode2 event bus

rxode2 has a small publish/subscribe event bus (see `?rxode2::rxEventListen`).
Packages in the nlmixr2 ecosystem emit an event when a user-facing operation
finishes (a fit, a simulation, a result), and any package can register a
listener to be told about it.  The main listener is
[nlmixr2log](https://github.com/nlmixr2/nlmixr2log), which records each fit
as a run, but you can write your own.

Events are delivered only from the outermost operation: while any operation
is running (an rxode2 "scope" is open), nested work emits nothing.  The
scope depth is mirrored in the `RXODE2_EVENT_DEPTH` environment variable,
so worker processes started inside an operation stay silent too.

`nlmixr2targets` emits events only when rxode2 has the bus; with an older
rxode2, and when nobody listens, nothing changes.

# What a `tar_nlmixr()` pipeline emits

`tar_nlmixr(name = <name>, ...)` generates four targets (see
`vignette("caching")`).  Two of them produce events:

| Target | Event | Emitted by |
| ------ | ----- | ---------- |
| `<name>_fit_simple` | `fitComplete` | nlmixr2est, from the `nlmixr2()` call that estimates the simplified model |
| `<name>` | `fitUpdate`, then `assign` | nlmixr2targets, after restoring the labels and the original data |

The `fitComplete` event comes from nlmixr2est itself (when its version has
the event hooks); `nlmixr2targets` only makes sure that a logger is running
in the process that estimates the model (see below).  The final `<name>`
target then reports the finished fit.

## `fitUpdate`

The final target restores parameter labels and the original data onto the
estimated fit in place.  It reports that change as a new version of the same
run:

| Field | Value |
| ----- | ----- |
| `fit` | the final fit (labels and original data restored) |
| `original` | the same fit (`original = fit`), so a logger treats it as the same run |
| `name` | `NULL` |
| `what` | `"ui"` |
| `inPlace` | `TRUE` |

## `assign`

Right after `fitUpdate`, the final target reports the name the fit is known
by, which is the target name:

| Field | Value |
| ----- | ----- |
| `name` | `targets::tar_name()`, i.e. `"<name>"` |
| `value` | the final fit |
| `call` | `NULL` |
| `envir` | `NULL` |
| `cached` | `FALSE` |

## One run per model

Taken together, a logger such as nlmixr2log sees, for each `tar_nlmixr()`
model:

1. `fitComplete` when `<name>_fit_simple` estimates the model (a new run),
2. `fitUpdate` with `inPlace = TRUE` and `original = fit` (a new version of
   that same run, not a new run), and
3. `assign` with the target name (the run is named `<name>`).

So each `tar_nlmixr()` model is stored as one run, named after its target,
rather than as one run for the simplified fit and another for the final fit.
For `tar_nlmixr_multimodel()`, each model's final target is named
`<name>_<hash>`, and that is the name the `assign` event carries.

## What is not emitted

* **Nothing from the other targets.**  `<name>_object_simple` and
  `<name>_data_simple` do no fitting and emit nothing.
* **No `assign` outside a pipeline.**  `assign` is emitted only while a
  target is running (`targets::tar_active()`).  Calling
  `nlmixr_object_complicate()` by hand emits only `fitUpdate`.
* **Nothing for failed fits.**  With `error = "continue"`, a failed
  estimation stores an `nlmixr2targetsError` object instead of a fit, and
  the final target emits nothing for it.
* **No `fitComplete` for a cached fit.**  When only cosmetic parts of the
  model change (such as parameter labels), `<name>_fit_simple` is not rerun
  (see `vignette("caching")`), so only the final target's `fitUpdate` and
  `assign` are emitted.
  When nothing changes, `tar_make()` runs no target and emits nothing.
* **Nothing nested.**  Work done inside the fit (solves, inner fits) is
  silenced by the bus depth rule.

# Logging on target workers

The commands that `tar_nlmixr()` generates load only `nlmixr2targets` and
`nlmixr2est`, so a fresh worker (for example with `tar_make()` running
targets in a `callr` process, or with `crew` workers) would never load a
logger, and its fits would go unrecorded.

To handle this, the estimating target and the final target load
`nlmixr2log` when the project has an nlmixr2log configuration, that is when
either

* the `NLMIXR2LOG_CONFIG` environment variable is set, or
* a `_nlmixr2log.rds` file exists in the project directory (the working
  directory of the target).

Save that configuration once with `nlmixr2log::nlmixr2logSaveConfig()`;
loading nlmixr2log on a worker then starts the configured loggers before
the fit, so it is logged on any worker.  Without a configuration, or when
nlmixr2log is not installed, nothing is loaded and nothing happens.
nlmixr2log is not a dependency of `nlmixr2targets`.

```{r, eval = FALSE}
# once, in the project directory
library(nlmixr2log)
# ... configure the loggers you want, then:
nlmixr2logSaveConfig()   # writes _nlmixr2log.rds

# afterwards, every tar_make() logs each tar_nlmixr() model as one run
targets::tar_make()
```

# Listening yourself

A listener is a function `function(event, ...)` that receives the event
name and the payload fields as named arguments.  This listener records what
it is told:

```{r listen}
recorded <- list()
rxode2::rxEventListen("demo", function(event, ...) {
  p <- list(...)
  recorded[[length(recorded) + 1L]] <<- list(
    event = event,
    name = p$name,
    what = p$what,
    inPlace = p$inPlace
  )
})
```

Run a small pipeline in a temporary directory.  `callr_function = NULL` runs
the targets in this R session, so the listener above hears them (a separate
worker process would need its own listener, which is what the nlmixr2log
configuration above provides).

```{r pipeline, message = FALSE, warning = FALSE, results = "hide"}
targets::tar_dir({
  targets::tar_script({
    library(nlmixr2targets)
    pheno <- function() {
      ini({
        lcl <- log(0.008); label("Typical value of clearance")
        lvc <- log(0.6); label("Typical value of volume of distribution")
        etalcl ~ 1
        cpaddSd <- 0.1
      })
      model({
        cl <- exp(lcl + etalcl)
        vc <- exp(lvc)
        kel <- cl / vc
        d / dt(central) <- -kel * central
        cp <- central / vc
        cp ~ add(cpaddSd)
      })
    }
    list(
      tar_nlmixr(
        name = pheno_model,
        object = pheno,
        data = nlmixr2data::pheno_sd,
        est = "posthoc"
      )
    )
  }, ask = FALSE)
  targets::tar_make(callr_function = NULL, reporter = "silent")
})
```

```{r show}
str(recorded)
```

With an nlmixr2est that has the event hooks, the list starts with a
`fitComplete` event from the `pheno_model_fit_simple` target; it always ends
with the `fitUpdate` (`what = "ui"`, `inPlace = TRUE`) and `assign`
(`name = "pheno_model"`) events from the final target.

Remove the listener when you are done:

```{r unlisten}
rxode2::rxEventUnlisten("demo")
```

For full logging (run storage, versions, names, and the files a run
produces), use [nlmixr2log](https://github.com/nlmixr2/nlmixr2log) instead of
writing your own listener.
