---
title: "Initial conditions cheatsheet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Initial conditions cheatsheet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nlmixr2targets)
```

# Two equivalent forms

nlmixr2's DSL accepts two notations for setting a compartment's initial
value inside a `model({...})` block. Both work in `tar_nlmixr()` and
`tar_nlmixr_multimodel()`, and both produce the same fit:

```{r both-forms, eval = FALSE}
# Form 1: natural DSL syntax
model({
  d/dt(central) <- -kel*central
  central(0) <- 0          # initial value at time 0
  cp <- central/vc
  cp ~ add(cpaddSd)
})

# Form 2: the older `cmt(initial) <- value` workaround
model({
  d/dt(central) <- -kel*central
  central(initial) <- 0    # same meaning as central(0) <- 0
  cp <- central/vc
  cp ~ add(cpaddSd)
})
```

# What `tar_nlmixr()` does behind the scenes

`targets` walks every function in the user's environment with
`codetools::findGlobals()` and rejects replacement-style assignments
whose target is not a symbol -- which is exactly what `central(0) <- 0`
looks like. To allow the natural DSL syntax, `tar_nlmixr()` walks the
captured `object` expression at construction time and rewrites every
`cmt(0) <- value` inside `model({...})` to `cmt(initial) <- value`. The
rewrite is reversed at runtime before the model is handed to nlmixr2, so
the fit you get is identical.

The rewrite mutates the function binding **in place** in the
environment where it was found, so if you later print

```{r print-body, eval = FALSE}
body(my_model)
```

at the REPL after calling `tar_nlmixr()`, you will see
`central(initial) <- 0` rather than the originally-typed
`central(0) <- 0`. This is the only user-visible side effect.

# Pipe forms

Both pipe forms below also work, including with `cmt(0)`:

```{r pipe-forms, eval = FALSE}
tar_nlmixr(
  name = m1,
  object = pheno |> model({ central(0) <- 0 }, append = TRUE),
  data  = nlmixr2data::pheno_sd,
  est   = "saem"
)

tar_nlmixr(
  name = m2,
  object = pheno |> ini(cpaddSd = 0.3),
  data  = nlmixr2data::pheno_sd,
  est   = "saem"
)
```

A runtime delayed-eval wrapper restores `cmt(0)` before nlmixr2's pipe
handlers see the parsed body.

# Known limitation: env functions never routed through `tar_nlmixr()`

`targets` walks **every** function in the analysed environment, not just
the ones referenced by a target. If you define

```{r unused-function, eval = FALSE}
sketch <- function() {
  ini({ a <- 1 })
  model({ central(0) <- 0 })
}
```

in the same environment but never pass `sketch` to `tar_nlmixr()`,
`targets`' static analysis will still try to walk its body and will
report a `codetools::findGlobals()` error. The workaround is either:

- pass the function through `tar_nlmixr()` (the construction-time
  rewriter cleans it up), or
- write the manual `central(initial) <- 0` form by hand.

This case is pinned by a test in `tests/testthat/test-tar_nlmixr.R`.

# Summary

- Both `cmt(0) <- value` and `cmt(initial) <- value` are accepted.
- `tar_nlmixr()` quietly rewrites the former to the latter in env, and
  back again at evaluation time.
- After `tar_nlmixr()`, `body(my_model)` will show the rewritten form;
  this is cosmetic only.
- Pipe forms work too via a runtime wrapper.
- Functions in env that contain `cmt(0)` but are never routed through
  `tar_nlmixr()` still need the manual `cmt(initial)` form.
